Solar-wind predictions for the Parker Solar Probe orbit

### 2017-11-20 paperMVVB_for_aanda/paperMVVB.tex
for aanda bib modified; for arXiv tex files modified

\begin{document}
        \title{Solar-wind predictions for the Parker Solar Probe orbit}
        \subtitle{Near-Sun extrapolations derived from an empirical solar-wind model based on Helios and OMNI observations}
        \author{M.~S.~Venzmer
        \and V.~Bothmer}
        \institute{University of Goettingen, Institute for Astrophysics, Friedrich-Hund-Platz~1, 37077~Göttingen, Germany}
        \date{Received 25 August 2017; accepted date}

@@

        From visual inspection, the resulting fit curves describe the shape of the magnetic field strength, the density and the temperature distributions well, as can be seen in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}.
        \begin{figure*}
                \includegraphics[width=18cm]{figures/histogram_fits_4_a_zoom_paper_pdfplot.pdf}
                \caption{Frequency distributions of the four solar-wind parameters and their lognormal fits derived from the hourly OMNI data set. The histograms have bins of \SI{0.5}{\nT}, \SI{10}{\km\per\s}, \SI{1}{\per\cm\cubed} and \SI{10000}{\K}. The fits' median and mean values are indicated as well. The insets show zoomed-in views of the high-value tails of the distributions.}
                \label{fig:histogram_fits_4_a_zoom_paper_pdfplot}
        \end{figure*}
        However, for the velocity, the fit function appears not to be as good in describing the measured distribution’s more complex shape around its peak and in the higher velocity range. This also can be inferred from the sum of absolute residuals (SAR) between data and fit, listed in Table~\ref{tab:lognormal_fit_parameters} as a percentage of the distribution area, being almost three times larger than those from the other parameters.
        In order to find a better fit result for the velocity distribution, we assume that the velocity distribution can be made up of at least two overlapping branches \citep{McGregor2011b}. Therefore a compositional approach  is chosen by combining two lognormal functions (\ref{eq:single_lognormal_fit_function}), involving more fit variables:
        \begin{align}
                W_\text{II}(x) &= c \cdot W_1(x) + (1 -c) \cdot W_2(x)\,.       \label{eq:double_lognormal_fit_function}
        \end{align}
        The balancing parameter $c$ ensures that the resulting function remains normalized as it represents a probability distribution.
        The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution yields the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$) as listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean values of the composed distribution, which can be derived by solving
        \begin{align}
                \int W_\text{II}(x)\,\text{d}x = 0      &       &\text{and}     &       &\int x\,W_\text{II}(x)\,\text{d}x = 0        \,.
        \end{align}
        This more complex fit function is more accurate in describing the velocity's frequency distribution as shown in Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
                \caption{Velocity frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset is a zoomed-in view of the high value tail of the distribution.}
                \label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
        \end{figure}
        Thus in the following Sections we keep the double lognormal ansatz for all velocity frequency fits.

@@

        Yearly medians of the solar-wind parameters and the yearly SSN together with the solar cycle number are shown in the upper part of Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot}. The reason for correlating the SSN to the solar-wind median values is because the position of a lognormal function is defined by its median. The data are averaged to yearly values to avoid seasonal effects during the Earth’s orbit around the Sun caused by its variations in solar latitude and distance.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
                \caption{Solar-wind parameter yearly frequencies (gray shading) with yearly medians (lines) derived from OMNI data and the yearly SSN from the \citet{sidc} with solar cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
                \label{fig:OMNI_yearly_ssn_correlation_c_plot}
        \end{figure}
        The solar-wind velocity, density, and temperature depend on the state of the solar cycle \citep{Schwenn1983}.
        For instance the fast solar wind occurs at times when polar coronal holes extend to lower latitudes, a typical feature of the declining phase of the solar cycle as pointed out by \citet[p.~75, Fig.~3.52]{Bothmer2007}. Therefore the solar-wind velocity, density, and temperature maxima exhibit time lags relative to the SSN maxima.

@@

        Naturally, the fit models match with the general data trends, as can be seen from Fig.~\ref{fig:OMNI_yearly_BVdblNTSSN_fit_e_plot}, though single year variations are not replicated by the model (e.g., the high velocity and temperature values in 1974, 1994, and 2003).
        \begin{figure*}
                \includegraphics[width=18cm]{figures/OMNI_yearly_BVdblNTSSN_fit_e_plot.pdf}
                \caption{Solar wind parameter yearly data frequencies and lognormal fit models, both with their median values (white lines) over the OMNI time period 1963--2016. The corresponding yearly SSN and the shifted SSN for the models are indicated by gray and black lines. The velocity median is derived from the SSN-weighted constant lognormal parts (dotted lines).}
                \label{fig:OMNI_yearly_BVdblNTSSN_fit_e_plot}
        \end{figure*}
        The comparison of this model with the yearly data median values with respect to the lagged SSN shows that the medians obtained from the modeling have a similar slope, as shown in Fig.~\ref{fig:OMNI_yearly_BVNTvsSSN_a}.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_BVNTvsSSN_a.pdf}}
                \caption{Solar-wind parameter medians with respect to the lagged SSN. The yearly data medians (+) with their weighted linear fit (solid lines) are obtained from OMNI data. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The SSN-dependent median (dashed lines) is derived from the lognormal model fit. For the velocity the median is derived from the SSN-weighting (\ref{eq:balance_with_ssn}) of the slow and fast model parts (dotted lines), whose magnitudes are SSN independent.}
                \label{fig:OMNI_yearly_BVNTvsSSN_a}
        \end{figure}

        Again, the solar-wind velocity needs a special treatment because of the application of the double lognormal distribution (\ref{eq:double_lognormal_fit_function}). Since it is well known that slow and fast solar-wind stream occurrence rates follow the solar cycle, we keep the two velocity components' positions SSN-independent ($x_\text{med} =  b_\text{med}$) and vary instead their balance with the SSN:
        \begin{align}
                c(ssn) &= c_a \cdot ssn + c_b\,.        \label{eq:balance_with_ssn}
        \end{align}
        The fit result (see Table~\ref{tab:ssn_fit_parameters}) yields a model in which three years after solar cycle minimum (SSN of zero) the contribution of slow solar wind to the overall solar wind distribution reaches a maximum value (about \SI{64}{\percent}) and decreases with increasing SSN as shown in Fig.~\ref{fig:Vdbl_SSN_ratio_f_plot}.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{figures/Vdbl_SSN_ratio_f_plot.pdf}}
                \caption{Ratio of slow to fast solar wind for a SSN lagged by 3~years. The yearly ratios (+) and their weighted linear fit (solid line) are obtained from OMNI data with a threshold velocity of $v_\text{th} = \SI{400}{\km\per\s}$. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The model's balance parameter (\ref{eq:balance_with_ssn}) and derived ratio (same threshold) are plotted as dashed and dotted lines.}
                \label{fig:Vdbl_SSN_ratio_f_plot}
        \end{figure}

@@

                \includegraphics[width=18cm]{figures/radial_fit_4_thesis_light_b_skip.pdf}

@@

                \includegraphics[width=18cm]{figures/mixed_fit_fixed_4_paper_f_plot.pdf}

@@

        This solar-wind model for the radial distance dependence is representative for the time of the Helios observations around the rise of solar cycle~21. The variations of the yearly power law fit exponents from fitting the solar-distance dependency (\ref{eq:power_function}) are shown in Fig.~\ref{fig:yearly_gradients_c} together with the yearly SSN for the time period \numrange{1974}{1982}. It can be seen that during the Helios time period there might be some systematic  variation of the exponents with solar activity -- at least for the velocity and temperature exponents.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{figures/yearly_gradients_c.pdf}}
                \caption{Helios yearly variation of the solar wind parameter power exponents for the dependence on radial distance together with the SIDC 13-month smoothed monthly SSN. The weighted standard deviations and average values for all years are indicated by the shaded areas. In this plot, the 21~days since Helios launch in the year 1974 are omitted because a distance range of merely \SIrange{0.95}{0.98}{\au} was covered that year.}
                \label{fig:yearly_gradients_c}
        \end{figure}
        However, for simplicity we assume that the distance scaling laws can be treated as time independent and include the calculated exponents’ yearly variations $\Delta e$, summarized in Table~\ref{tab:mean_median_fit_parameter}, as relative uncertainties.

@@

        Parker Solar Probe is planned to launch in mid 2018. With its first Venus flyby it will swing into Venus' orbital plane, reaching  a first perihelion with a distance of \SI{0.16}{\au  } just 93~days
after launch, in November 2018. Seven additional Venus flybys allow the perihelion distance to be reduced to a minimum of \SI{9.86}{\Rs}. This distance will be reached with the 22nd perihelion in December 2024 \citep{Fox2015}, as plotted in the top panel of Fig.~\ref{fig:SPP_orbit_predicted_SSN_overview_f_plot}.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{figures/SPP_orbit_predicted_SSN_overview_f_plot.pdf}}
                \caption{PSP's solar distance during its mission time (top). Consecutive Venus flybys bring its perihelia nearer to the Sun. Actual and predicted SSN (bottom), that is, SIDC 13-month smoothed monthly actual SSN, SIDC Standard Curves Kalman filter prediction and SWPC prediction with their corresponding expected ranges (shaded areas). The SSN from previous cycle~24, shifted by 11~years, is plotted together with two alternative trends of half and twice its amplitude.}
                \label{fig:SPP_orbit_predicted_SSN_overview_f_plot}
        \end{figure}

@@

                \includegraphics[width=18cm]{figures/sw_extrapolation_ssn_f_plot.pdf}

@@

        Implementing the SSN predictions for the PSP mission time and the orbital trajectory data, we can infer which solar-wind parameter magnitudes can be expected. Figures~\ref{fig:SPP_perihelia_prediction_f_plot} and \ref{fig:SPP_perihelia_prediction_nearest_f_plot} show the median values (\ref{eq:median_with_ssn}) of the considered different solar-wind parameters for 12-day periods, comprising the first perihelion in Novemver 2018 and the first closest perihelion in December 2024.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_f_plot.pdf}}
                \caption{Estimated solar-wind parameter medians (black lines) and their error bands (gray) during 12~days in 2018 with PSP's first perihelion at about \SI{0.16}{\au}. For the velocity the combined median is calculated and also the SSN-independent slow and fast parts are plotted (dotted lines).}
                \label{fig:SPP_perihelia_prediction_f_plot}
        \end{figure}
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_nearest_f_plot.pdf}}
                \caption{Estimated solar-wind parameter medians (black lines) and their error bands (gray) during 12~days in 2024 with PSP's 22nd (the first closest) perihelion at \SI{0.0459}{\au}. For the velocity the combined median is calculated and also the SSN-independent slow and fast parts are plotted (dotted lines).}
                \label{fig:SPP_perihelia_prediction_nearest_f_plot}
        \end{figure}
        In the beginning of the mission median values of about \SI{87}{\nT}, \SI{340}{\km\per\s}, \SI{214}{\per\cm\cubed} and \SI{503000}{\K} are estimated to be measured at \SI{0.16}{\au}, increasing to about \SI{943}{\nT}, \SI{290}{\km\per\s}, \SI{2951}{\per\cm\cubed} and \SI{1930000}{\K} during the first closest approach at \SI{0.046}{\au}. Monthly SSNs -- shifted by the time lags specific to the solar-wind parameters -- are used in the calculation of the solar-wind predictions. These SSNs are either actual smoothed values from the SIDC with their reported standard deviations, short-term predictions from the SWPC with their expected ranges, or actual smoothed values from the SIDC shifted by 11~years with half/twice their values as uncertainties. The error bands given in both Figures, calculated from error propagation, include these SSN ranges and the derived fit parameter errors.

        Finally the estimated solar-wind environment can be derived from the function $W'''(x,ssn,r)$. The estimated frequency distributions of the four solar-wind parameters at PSP's 1st and 22nd (first closest) perihelion are plotted in Fig.~\ref{fig:SPP_sw_distributions_b}.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{figures/SPP_sw_distributions_b.pdf}}
                \caption{Frequency distributions of the four solar-wind parameters (same as in Figs.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot} and \ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}) and those estimated with the solar-wind model for PSP's 1st and 22nd (first closest) perihelion. In these Figures the frequencies of both extrapolated curves are scaled for visibility to the same height as the \SI{1}{\au} distribution.}
                \label{fig:SPP_sw_distributions_b}
        \end{figure}
        Again, we point out that the velocity and temperature distributions for the 22nd perihelion are only upper limits and the actual values to be encountered by PSP are expected to be smaller.

>>>>

\begin{document}
        \title{Solar-wind predictions for the Parker Solar Probe orbit}
        \subtitle{Near-Sun extrapolations derived from an empirical solar-wind model based on Helios and OMNI observations}
        \author{M.~S.~Venzmer
        \and V.~Bothmer}
        \institute{University of Goettingen, Institute for Astrophysics, Friedrich-Hund-Platz~1, 37077~Göttingen, Germany}
        \date{Received 25 August 2017; accepted 10 November 2017}

@@

        From visual inspection, the resulting fit curves describe the shape of the magnetic field strength, the density and the temperature distributions well, as can be seen in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}.
        \begin{figure*}
                \includegraphics[width=18cm]{histogram_fits_4_a_zoom_paper_pdfplot.pdf}
                \caption{Frequency distributions of the four solar-wind parameters and their lognormal fits derived from the hourly OMNI data set. The histograms have bins of \SI{0.5}{\nT}, \SI{10}{\km\per\s}, \SI{1}{\per\cm\cubed} and \SI{10000}{\K}. The fits' median and mean values are indicated as well. The insets show zoomed-in views of the high-value tails of the distributions.}
                \label{fig:histogram_fits_4_a_zoom_paper_pdfplot}
        \end{figure*}
        However, for the velocity, the fit function appears not to be as good in describing the measured distribution’s more complex shape around its peak and in the higher velocity range. This also can be inferred from the sum of absolute residuals (SAR) between data and fit, listed in Table~\ref{tab:lognormal_fit_parameters} as a percentage of the distribution area, being almost three times larger than those from the other parameters.
        In order to find a better fit result for the velocity distribution, we assume that the velocity distribution can be made up of at least two overlapping branches \citep{McGregor2011b}. Therefore a compositional approach  is chosen by combining two lognormal functions (\ref{eq:single_lognormal_fit_function}), involving more fit variables:
        \begin{align}
                W_\text{II}(x) &= c \cdot W_1(x) + (1 -c) \cdot W_2(x)\,.       \label{eq:double_lognormal_fit_function}
        \end{align}
        The balancing parameter $c$ ensures that the resulting function remains normalized as it represents a probability distribution.
        The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution yields the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$) as listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean values of the composed distribution, which can be derived by solving
        \begin{align}
                \int W_\text{II}(x)\,\text{d}x = 0      &       &\text{and}     &       &\int x\,W_\text{II}(x)\,\text{d}x = 0        \,.
        \end{align}
        This more complex fit function is more accurate in describing the velocity's frequency distribution as shown in Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
                \caption{Velocity frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset is a zoomed-in view of the high value tail of the distribution.}
                \label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
        \end{figure}
        Thus in the following Sections we keep the double lognormal ansatz for all velocity frequency fits.

@@

        Yearly medians of the solar-wind parameters and the yearly SSN together with the solar cycle number are shown in the upper part of Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot}. The reason for correlating the SSN to the solar-wind median values is because the position of a lognormal function is defined by its median. The data are averaged to yearly values to avoid seasonal effects during the Earth’s orbit around the Sun caused by its variations in solar latitude and distance.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{OMNI_yearly_ssn_correlation_c_plot.pdf}}
                \caption{Solar-wind parameter yearly frequencies (gray shading) with yearly medians (lines) derived from OMNI data and the yearly SSN from the \citet{sidc} with solar cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
                \label{fig:OMNI_yearly_ssn_correlation_c_plot}
        \end{figure}
        The solar-wind velocity, density, and temperature depend on the state of the solar cycle \citep{Schwenn1983}.
        For instance the fast solar wind occurs at times when polar coronal holes extend to lower latitudes, a typical feature of the declining phase of the solar cycle as pointed out by \citet[p.~75, Fig.~3.52]{Bothmer2007}. Therefore the solar-wind velocity, density, and temperature maxima exhibit time lags relative to the SSN maxima.

@@

        Naturally, the fit models match with the general data trends, as can be seen from Fig.~\ref{fig:OMNI_yearly_BVdblNTSSN_fit_e_plot}, though single year variations are not replicated by the model (e.g., the high velocity and temperature values in 1974, 1994, and 2003).
        \begin{figure*}
                \includegraphics[width=18cm]{OMNI_yearly_BVdblNTSSN_fit_e_plot.pdf}
                \caption{Solar wind parameter yearly data frequencies and lognormal fit models, both with their median values (white lines) over the OMNI time period 1963--2016. The corresponding yearly SSN and the shifted SSN for the models are indicated by gray and black lines. The velocity median is derived from the SSN-weighted constant lognormal parts (dotted lines).}
                \label{fig:OMNI_yearly_BVdblNTSSN_fit_e_plot}
        \end{figure*}
        The comparison of this model with the yearly data median values with respect to the lagged SSN shows that the medians obtained from the modeling have a similar slope, as shown in Fig.~\ref{fig:OMNI_yearly_BVNTvsSSN_a}.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{OMNI_yearly_BVNTvsSSN_a.pdf}}
                \caption{Solar-wind parameter medians with respect to the lagged SSN. The yearly data medians (+) with their weighted linear fit (solid lines) are obtained from OMNI data. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The SSN-dependent median (dashed lines) is derived from the lognormal model fit. For the velocity the median is derived from the SSN-weighting (\ref{eq:balance_with_ssn}) of the slow and fast model parts (dotted lines), whose magnitudes are SSN independent.}
                \label{fig:OMNI_yearly_BVNTvsSSN_a}
        \end{figure}

        Again, the solar-wind velocity needs a special treatment because of the application of the double lognormal distribution (\ref{eq:double_lognormal_fit_function}). Since it is well known that slow and fast solar-wind stream occurrence rates follow the solar cycle, we keep the two velocity components' positions SSN-independent ($x_\text{med} =  b_\text{med}$) and vary instead their balance with the SSN:
        \begin{align}
                c(ssn) &= c_a \cdot ssn + c_b\,.        \label{eq:balance_with_ssn}
        \end{align}
        The fit result (see Table~\ref{tab:ssn_fit_parameters}) yields a model in which three years after solar cycle minimum (SSN of zero) the contribution of slow solar wind to the overall solar wind distribution reaches a maximum value (about \SI{64}{\percent}) and decreases with increasing SSN as shown in Fig.~\ref{fig:Vdbl_SSN_ratio_f_plot}.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{Vdbl_SSN_ratio_f_plot.pdf}}
                \caption{Ratio of slow to fast solar wind for a SSN lagged by 3~years. The yearly ratios (+) and their weighted linear fit (solid line) are obtained from OMNI data with a threshold velocity of $v_\text{th} = \SI{400}{\km\per\s}$. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The model's balance parameter (\ref{eq:balance_with_ssn}) and derived ratio (same threshold) are plotted as dashed and dotted lines.}
                \label{fig:Vdbl_SSN_ratio_f_plot}
        \end{figure}

@@

                \includegraphics[width=18cm]{radial_fit_4_thesis_light_b_skip.pdf}

@@

                \includegraphics[width=18cm]{mixed_fit_fixed_4_paper_f_plot.pdf}

@@

        This solar-wind model for the radial distance dependence is representative for the time of the Helios observations around the rise of solar cycle~21. The variations of the yearly power law fit exponents from fitting the solar-distance dependency (\ref{eq:power_function}) are shown in Fig.~\ref{fig:yearly_gradients_c} together with the yearly SSN for the time period \numrange{1974}{1982}. It can be seen that during the Helios time period there might be some systematic  variation of the exponents with solar activity -- at least for the velocity and temperature exponents.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{yearly_gradients_c.pdf}}
                \caption{Helios yearly variation of the solar wind parameter power exponents for the dependence on radial distance together with the SIDC 13-month smoothed monthly SSN. The weighted standard deviations and average values for all years are indicated by the shaded areas. In this plot, the 21~days since Helios launch in the year 1974 are omitted because a distance range of merely \SIrange{0.95}{0.98}{\au} was covered that year.}
                \label{fig:yearly_gradients_c}
        \end{figure}
        However, for simplicity we assume that the distance scaling laws can be treated as time independent and include the calculated exponents’ yearly variations $\Delta e$, summarized in Table~\ref{tab:mean_median_fit_parameter}, as relative uncertainties.

@@

        Parker Solar Probe is planned to launch in mid 2018. With its first Venus flyby it will swing into Venus' orbital plane, reaching  a first perihelion with a distance of \SI{0.16}{\au  } just 93~days
after launch, in November 2018. Seven additional Venus flybys allow the perihelion distance to be reduced to a minimum of \SI{9.86}{\Rs}. This distance will be reached with the 22nd perihelion in December 2024 \citep{Fox2015}, as plotted in the top panel of Fig.~\ref{fig:SPP_orbit_predicted_SSN_overview_f_plot}.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{SPP_orbit_predicted_SSN_overview_f_plot.pdf}}
                \caption{PSP's solar distance during its mission time (top). Consecutive Venus flybys bring its perihelia nearer to the Sun. Actual and predicted SSN (bottom), that is, SIDC 13-month smoothed monthly actual SSN, SIDC Standard Curves Kalman filter prediction and SWPC prediction with their corresponding expected ranges (shaded areas). The SSN from previous cycle~24, shifted by 11~years, is plotted together with two alternative trends of half and twice its amplitude.}
                \label{fig:SPP_orbit_predicted_SSN_overview_f_plot}
        \end{figure}

@@

                \includegraphics[width=18cm]{sw_extrapolation_ssn_f_plot.pdf}

@@

        Implementing the SSN predictions for the PSP mission time and the orbital trajectory data, we can infer which solar-wind parameter magnitudes can be expected. Figures~\ref{fig:SPP_perihelia_prediction_f_plot} and \ref{fig:SPP_perihelia_prediction_nearest_f_plot} show the median values (\ref{eq:median_with_ssn}) of the considered different solar-wind parameters for 12-day periods, comprising the first perihelion in Novemver 2018 and the first closest perihelion in December 2024.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{SPP_perihelia_prediction_f_plot.pdf}}
                \caption{Estimated solar-wind parameter medians (black lines) and their error bands (gray) during 12~days in 2018 with PSP's first perihelion at about \SI{0.16}{\au}. For the velocity the combined median is calculated and also the SSN-independent slow and fast parts are plotted (dotted lines).}
                \label{fig:SPP_perihelia_prediction_f_plot}
        \end{figure}
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{SPP_perihelia_prediction_nearest_f_plot.pdf}}
                \caption{Estimated solar-wind parameter medians (black lines) and their error bands (gray) during 12~days in 2024 with PSP's 22nd (the first closest) perihelion at \SI{0.0459}{\au}. For the velocity the combined median is calculated and also the SSN-independent slow and fast parts are plotted (dotted lines).}
                \label{fig:SPP_perihelia_prediction_nearest_f_plot}
        \end{figure}
        In the beginning of the mission median values of about \SI{87}{\nT}, \SI{340}{\km\per\s}, \SI{214}{\per\cm\cubed} and \SI{503000}{\K} are estimated to be measured at \SI{0.16}{\au}, increasing to about \SI{943}{\nT}, \SI{290}{\km\per\s}, \SI{2951}{\per\cm\cubed} and \SI{1930000}{\K} during the first closest approach at \SI{0.046}{\au}. Monthly SSNs -- shifted by the time lags specific to the solar-wind parameters -- are used in the calculation of the solar-wind predictions. These SSNs are either actual smoothed values from the SIDC with their reported standard deviations, short-term predictions from the SWPC with their expected ranges, or actual smoothed values from the SIDC shifted by 11~years with half/twice their values as uncertainties. The error bands given in both Figures, calculated from error propagation, include these SSN ranges and the derived fit parameter errors.

        Finally the estimated solar-wind environment can be derived from the function $W'''(x,ssn,r)$. The estimated frequency distributions of the four solar-wind parameters at PSP's 1st and 22nd (first closest) perihelion are plotted in Fig.~\ref{fig:SPP_sw_distributions_b}.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{SPP_sw_distributions_b.pdf}}
                \caption{Frequency distributions of the four solar-wind parameters (same as in Figs.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot} and \ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}) and those estimated with the solar-wind model for PSP's 1st and 22nd (first closest) perihelion. In these Figures the frequencies of both extrapolated curves are scaled for visibility to the same height as the \SI{1}{\au} distribution.}
                \label{fig:SPP_sw_distributions_b}
        \end{figure}
        Again, we point out that the velocity and temperature distributions for the 22nd perihelion are only upper limits and the actual values to be encountered by PSP are expected to be smaller.
